\chapter*{Abstract}
\paragraph{}

The intracellular signalling network involved in apoptosis is a complex system crucial to regulating cell death. Despite the advancements made in the field of cell signalling, there remain several limitations in our understanding of these pathways. The study focused on analyzing the interactions between the STAT1, STAT3, Bcl-2, and BAX proteins and their impact on the network. This research contributes to the advancement of our understanding of the intracellular signalling network involved in apoptosis and highlights the importance of mathematical modelling approaches in the study of cell signalling. This study focuses on mathematical modeling approaches for cell signalling and aims to understand the genetic alterations in signalling pathways in cancer and their downstream signalling. The project explores the intracellular signalling network of proteins STAT1, STAT3, Bcl-2, and BAX and presents a mathematical model of the apoptosis signalling network. The mathematical model includes ordinary differential equations that govern the intracellular signalling network, and the numerical results were obtained using Euler's method and Runge-Kutta fourth-order method. The results of the study show that there are three equilibrium points when S = 0.4, two of which are stable with AB $<$ 1, while the third is unstable with AB $>$ 1. The results also show that when S = 0, the equilibrium point is stable, and when S = 1, the equilibrium point is also stable with AB $<$ 1. The study concludes with a summary of the findings and recommendations for future research. This study provides valuable insights into the mathematical modeling of cell signalling and how genetic alterations in signalling pathways can impact downstream signalling in cancer. It highlights the importance of mathematical modeling in understanding complex biological systems and provides a foundation for further research in this field.
