\chapter{Conclusion}
\label{chap:06}

\paragraph{}

Apoptosis is a crucial biological process that regulates cell death in response to various stimuli. It is a complex process that involves several key signalling pathways and inter-cellular modules. The Intercellular modules such as STAT1, STAT3, Bcl-2 and BAX play a critical role in apoptosis signalling. Understanding the dynamics of these modules and how they interact with each other is crucial for gaining insights into the apoptosis process. The project focused on two methods for numerical investigations, namely Euler’s method and the Runge-Kutta fourth-order method. The main focus of this project was to understand the behaviour of these modules in response to the activation of the IFN-$\beta$ and JAK2 signalling pathways. 

The numerical investigations were conducted by representing the levels of the Inter-cellular modules (STAT1, STAT3, Bcl-2, and BAX) and their activity by mathematical variables. The behaviour of the modules was analyzed and compared using both Euler’s method and the Runge-Kutta fourth-order method. The results of the numerical investigations were visualized and compared, allowing for a better understanding of the four behaviour of the Intercellular modules in response to the activation of the IFN-$\beta$ and JAK2
signalling pathways. 

In conclusion, the project will provide a comprehensive understanding of the behaviour of Intercellular modules involved in apoptosis signalling by combining the mathematical equilibrium and stability analysis and numerical investigations. The results will contribute to the field of apoptosis research and provide a foundation for further studies on the apoptosis signalling network.